---
title: "An analysis of HiChIP data with HiCDCPlus"
author: "Tim Myers"
date: "12/14/2021"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

This is a HiChIP analysis using the HiCDCPlus `R` package.  The input files used came from the output of a HiC-Pro analysis using its included sample data.  The input files are located in the /data folder.

Here I adapt the workflow outlined on the HiCDCPlus github page:
(https://github.com/mervesa/HiCDCPlus#ice-normalization-using-hitc-ice).

I identify significant interactions and differential interactions.

## Finding significant interactions from HiChIP

Load required libraries
```{r load-lib, eval=TRUE}
library(HiCDCPlus)
library(BSgenome.Hsapiens.UCSC.hg19)
library(DESeq2)
# library(HiTC)
```

# Generate Features
```{r features, eval=TRUE}
# define output directory path
out_dir <- "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/Output"

# generate genomic features
construct_features(output_path = paste0(out_dir, "/hg19_5kb_hindiii"),
                   gen = "Hsapiens",
                   gen_ver = "hg19",
                   sig = "AGCTT",
                   bin_type = "Bins-uniform",
                   binsize = 5000,
                   chrs = c("chr21","chr22")
                  )
```

Generate `gi_list` to store counts and features.
```{r gen-gi-list, eval = TRUE}

# add HiCPro .allValidPairs counts
### define path to .allValidPairs files ###

indexfile<-data.frame()

allValidPairsfile_paths <- c(
        # load each twice to simulate biological repleates required for differential interaction analysis
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results/data/dixon_2M/dixon_2M_rep1.allValidPairs",
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results/data/dixon_2M/dixon_2M_rep2.allValidPairs",
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results/data/dixon_2M_2/dixon_2M_2_rep1.allValidPairs",
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results/data/dixon_2M_2/dixon_2M_2_rep2.allValidPairs"
     )
      
for(allValidPairsfile_path in allValidPairsfile_paths){
  output_path <- paste0(out_dir,'/',
                        gsub("^(.*[\\/])", 
                             "",
                             gsub('.hic', '.txt.gz', allValidPairsfile_path)
                            )
                       )
  # generate gi_list instance
  gi_list <- generate_bintolen_gi_list(bintolen_path = paste0(out_dir,"/hg19_5kb_hindiii_bintolen.txt.gz"),
                                       gen = "Hsapiens", gen_ver="hg19"
                                      )
  gi_list <- add_hicpro_allvalidpairs_counts(gi_list, allvalidpairs_path = allValidPairsfile_path)

  # expand features for modeling
  gi_list <- expand_1D_features(gi_list)

  # run HiC-DC+ on 2 cores
  set.seed(1010) # HiC-DC down samples rows for modeling
  
  gi_list <- HiCDCPlus(gi_list,ssize = 0.1)
  for (i in seq(length(gi_list))){
    indexfile<-unique(rbind(indexfile,
                            as.data.frame(gi_list[[i]][gi_list[[i]]$qvalue<=0.05])[c('seqnames1',
                                                                                     'start1',
                                                                                     'start2')]
                            )
                      )
  }

  # write results to a text file
  gi_list_write(gi_list, fname = output_path)
}

```

Save the index file - union of significants at 5kb
``` {r save-index, eval = TRUE}

# change column names
colnames(indexfile) <-c ('chr', 'startI', 'startJ')

# write index file
data.table::fwrite(indexfile,
                   paste0(out_dir,'/dixon_analysis_indices.txt.gz'),
                   sep='\t',row.names=FALSE,quote=FALSE)

```

Next, get the union set of significant interactions and save it as the index file and then run `hicdcdiff`

