---
title: "An analysis of HiChIP data from Lyu et al. with HiCDCPlus"
author: "Tim Myers"
date: "01/05/2022"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

This is a HiChIP analysis using the HiCDCPlus `R` package.  The input files used came from the Lyu et al.(1) dataset containing HiChIP data derived from the immuno-precipitation of Rad21 in human embryonic stem cells (hESC) before and after heat shock.

Here use the workflow outlined in Chapter 11: Analysis of HiChIP Data in "Hi-C Data Analysis"[(2)](https://link.springer.com/protocol/10.1007/978-1-0716-1390-0_4).  Specifically, I only performed the initial steps of mapping, filtering and merging of the .fastq from Lyu et al. using HiC-Pro.  Then, using the .allValidPairs output filesare used as input for the HiCDCPlus downstream analysis. 

## Finding significant interactions from HiChIP

Load required libraries
```{r load-lib, eval=TRUE, message=FALSE}
library(HiCDCPlus)
library(BSgenome.Hsapiens.UCSC.hg19)
library(DESeq2)
# library(HiTC)
```

# Generate Features
```{r features, eval=TRUE}
# define output directory path
out_dir <- "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/Output_rad21"

# generate genomic features
construct_features(output_path = paste0(out_dir, "/hg19_50kb_dpnii"),
                   gen = "Hsapiens",
                   gen_ver = "hg19",
                   sig = "GATC", # DpnII enzyme, isoschizomer of MboI
                   bin_type = "Bins-uniform",
                   binsize = 50000,
                   chrs = c("chr21","chr22")
                  )
```

Generate `gi_list` to store counts and features.
```{r gen-gi-list, eval = TRUE}

# add HiCPro .allValidPairs counts
### define path to .allValidPairs files ###

indexfile<-data.frame()

allValidPairsfile_paths <- c(
        # load biological replicates for no treatment (NT) and heat-shock treated (HS)
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results_rad21/Rad21_HS_Rep1/Rad21_HS_Rep1.allValidPairs",
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results_rad21/Rad21_HS_Rep2/Rad21_HS_Rep2.allValidPairs",
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results_rad21/Rad21_NT_Rep1/Rad21_NT_Rep1.allValidPairs",
      "/Users/myersta/Box/Rprojects/hichip/hicpro_hicdcplus_test/data/hicpro_output/hic_results_rad21/Rad21_NT_Rep2/Rad21_NT_Rep2.allValidPairs"
     )
      
for(allValidPairsfile_path in allValidPairsfile_paths){
  
  output_path_hic <- paste0(out_dir,'/',
                            gsub("^(.*[\\/])", 
                                 "",
                                 gsub('.allValidPairs', '.hic', allValidPairsfile_path)
                                )
                            )
  
  output_path_results <- paste0(out_dir,'/',
                                gsub("^(.*[\\/])", 
                                     "",
                                     gsub('.allValidPairs', '.txt.gz', allValidPairsfile_path)
                                    )
                             )
  # generate gi_list instance
  gi_list <- generate_bintolen_gi_list(bintolen_path = paste0(out_dir,"/hg19_50kb_dpnii_bintolen.txt.gz"),
                                       gen = "Hsapiens", gen_ver="hg19"
                                      )
  
  # add .allValidPairs counts
  gi_list <- add_hicpro_allvalidpairs_counts(gi_list, allvalidpairs_path = allValidPairsfile_path)

  # expand features for modeling
  gi_list <- expand_1D_features(gi_list)

  # run HiC-DC+
  set.seed(1010) # HiC-DC down samples rows for modeling
  
  gi_list <- HiCDCPlus(gi_list,ssize = 0.1)
  for (i in seq(length(gi_list))){
    indexfile<-unique(rbind(indexfile,
                            as.data.frame(gi_list[[i]][gi_list[[i]]$qvalue<=0.05])[c('seqnames1',
                                                                                     'start1',
                                                                                     'start2')]
                            )
                      )
  }

  # write normalized counts (observed/expected) to a .hic file
  hicdc2hic(gi_list, hicfile = output_path_hic,
          mode='normcounts',gen_ver='hg19')
  
  # write results to a text file
  gi_list_write(gi_list, fname = output_path_results)
}

```

Save the index file - union of significants at 50kb
``` {r save-index, eval = TRUE}

# change column names
colnames(indexfile) <-c ('chr', 'startI', 'startJ')

# write index file
data.table::fwrite(indexfile,
                   paste0(out_dir,'/rad21_analysis_indices.txt.gz'),
                   sep='\t',row.names=FALSE,quote=FALSE)


```

Next, get the union set of significant interactions and save it as the index file and then run `hicdcdiff`
```{r get-union, eval = TRUE}

# Differential analysis using modified DESeq2
hicdcdiff(input_paths = list(rad21_HS = c(paste0(out_dir, '/Rad21_HS_Rep1.txt.gz'),
                                          paste0(out_dir, '/Rad21_HS_Rep2.txt.gz')),
                             rad21_NT = c(paste0(out_dir, '/Rad21_NT_Rep1.txt.gz'),
                                          paste0(out_dir, '/Rad21_NT_Rep2.txt.gz'))),
          filter_file = paste0(out_dir,'/rad21_analysis_indices.txt.gz'),
          output_path = paste0(out_dir,'/rad21_diff_analysis/'),
          fitType = 'mean',
          binsize = 50000,
          diagnostics = TRUE
         )
```

Analysis completed without errors.

```{r session, echo = TRUE}
writeLines(capture.output(sessionInfo()), paste0(out_dir, '/rad21_sessionInfo.txt'))
```

## References
1 Lyu X, Rowley MJ, Corces VG. Architectural proteins and pluripotency factors cooperate to orchestrate the transcriptional response of hESCs to temperature stress. Molecular cell. 2018 Sep 20;71(6):940-55.

2 Nicoletti C. Methods for the Differential Analysis of Hi-C Data. InHi-C Data Analysis 2022 (pp. 61-95). Humana, New York, NY.
